---
title: "MGT-6203 Project Idea - Jeff Hedberg"
date: 2022-06-16
output: 
  html_document:
    theme:
      version: 4
---

<style type="text/css">
.main-container {
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/GA Tech/MGT-6203/Project/vehicle_data")
```
<br/>  

### Project Idea - Auto Collision Repair Center  
##### An auto collision repair center in NYC needs a method for forcasting the volume of vehicles they will need to repair.  
##### This is very important since COVID-19 has had a large effect on their car repair volume.  
<br/>  

##### Steps to solve this problem:  
###### 1. Combine data from NYC Collisions, Vehicles and Persons datasets  
###### 2. Reshape data and compute necessary features for developing vehicle repair volume models   
###### 3. Create train test splits for modleing  
###### 4. Build vehicle repair volume models (several types)  
###### 5. Select the best performing model and generate predicitons for future year (assuming market share percentage)  
###### 6. compare Modeled accuracy vs actual accuracy   
###### 7. (Optional) Estimate effect of advertising campaign to increase market share and compute results  
###### &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; (1) What are the key demographics to target?   
###### &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; (2) How will this be delivered?    
###### &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; (3) How much will this cost?   
###### &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; (4) What will be the resulting increase in vehicle repair volume?  

<br/>  
<br/>  


##### Import Libraries
```{r, message=FALSE, warning=FALSE}
#### Load libraries
library(dplyr) # Used for data manipulation
library(knitr) # Used for R-Markdown knitting
library(kableExtra) # Used for Kable Tables
library(plotly) # Used for plotting
library(lubridate) #Used for Date manipulation
library(randomForest) # Used for building Random Forest models
library(rpart) # Used for building CART models
library(rpart.plot) # Used for plotting tree
```
<br/>  

##### Data Import, Raw Summaries  
##### Crashes Dataset
```{r, fig.width = 10, cache=FALSE}
#### Load Data
crashes_df <- read.csv('./Motor_Vehicle_Collisions_-_Crashes.csv', stringsAsFactors = FALSE) %>% 
  mutate(CRASH.DATE = as.Date(CRASH.DATE, "%m/%d/%Y")) #1,896,229 x 29

# crashes_df
# min(crashes_df$CRASH.DATE) #"2012-07-01"
# max(crashes_df$CRASH.DATE) #"2022-05-29

kable(t(summary(crashes_df))) %>% kable_classic(full_width = TRUE, html_font = "Cambria", font_size = 14)
```
<br/>  

##### Person Dataset
```{r, fig.width = 10, cache=FALSE}
#### Load Data
person_df <- read.csv('./Motor_Vehicle_Collisions_-_Person.csv', stringsAsFactors = FALSE) %>%
  mutate(CRASH_DATE = as.Date(CRASH_DATE, "%m/%d/%Y")) #4,692,054 x 21

# person_df
# min(person_df$CRASH_DATE) #"2012-07-01"
# max(person_df$CRASH_DATE) #"2022-05-29"

kable(t(summary(person_df))) %>% kable_classic(full_width = TRUE, html_font = "Cambria", font_size = 14)
```
<br/>  

##### Vehicle Dataset
```{r, fig.width = 10, cache=FALSE}
#### Load Data
vehicles_df <- read.csv('./Motor_Vehicle_Collisions_-_Vehicles.csv', stringsAsFactors = FALSE) %>%
  mutate(CRASH_DATE = as.Date(CRASH_DATE, "%m/%d/%Y")) #3,704,406 x 25

# vehicles_df
# min(vehicles_df$CRASH_DATE) #"2012-07-01"
# max(vehicles_df$CRASH_DATE) #"2021-12-04"

kable(t(summary(vehicles_df))) %>% kable_classic(full_width = TRUE, html_font = "Cambria", font_size = 14)
```
<br/>  


<br/>  
<br/>  

##### Show Issue with Person Dataset
```{r, fig.width = 10, cache=FALSE}
monthly_agg1 <- (person_df %>% 
                   mutate(yr_mo = paste0(substr(CRASH_DATE,1,4),'-',substr(CRASH_DATE,6,7))) %>% 
                   group_by(yr_mo) %>% 
                   summarise(n = n()) %>% 
                   arrange(yr_mo)
                 )

plot_ly(x=monthly_agg1$yr_mo, y=monthly_agg1$n, type='bar')
```
**Need to use adjust timeframe due to missing data.  Using data after 2017-01-01 for full year modeling**  
<br/>  


##### Process data to get to desired structure for Collision Volume Modeling
```{r, fig.width = 10, cache=FALSE}
combined_df <- (crashes_df %>% 
                  select(-c(CRASH.DATE, CRASH.TIME, CONTRIBUTING.FACTOR.VEHICLE.1, CONTRIBUTING.FACTOR.VEHICLE.2,
                            CONTRIBUTING.FACTOR.VEHICLE.3, CONTRIBUTING.FACTOR.VEHICLE.4, CONTRIBUTING.FACTOR.VEHICLE.5,
                            VEHICLE.TYPE.CODE.4, VEHICLE.TYPE.CODE.5,
                            NUMBER.OF.PEDESTRIANS.INJURED, NUMBER.OF.PEDESTRIANS.KILLED,
                            NUMBER.OF.CYCLIST.INJURED, NUMBER.OF.CYCLIST.KILLED,
                            ON.STREET.NAME, CROSS.STREET.NAME, OFF.STREET.NAME)) %>% 
                  inner_join(vehicles_df %>% 
                               filter(!is.na(VEHICLE_ID)) %>% 
                               select(-c(CRASH_DATE, CRASH_TIME, VEHICLE_ID))
                             , by = 'COLLISION_ID') %>% 
                  inner_join(person_df %>% 
                               select(-c(UNIQUE_ID, CONTRIBUTING_FACTOR_1, CONTRIBUTING_FACTOR_2))
                             , by= c('COLLISION_ID'='COLLISION_ID', 'UNIQUE_ID'='VEHICLE_ID')) %>% 
                  filter((PED_ROLE %in% c('Driver'))) %>%  #drivers only
                  filter((CRASH_DATE >= '2017-01-01') & (CRASH_DATE < '2021-12-01')) %>% #only from 2017-01-01 to 2021-11-30
                  filter(PERSON_AGE > 14 & PERSON_AGE< 101) %>% 
                  filter(PERSON_SEX == 'F' | PERSON_SEX=='M') %>% 
                  mutate(MONTH = substr(CRASH_DATE,6,7)) %>% 
                  mutate(TIMESTEP = as.period(interval(as.Date('2017-01-01'), CRASH_DATE)) %/% months(1)) %>% 
                  mutate(yr_mo = paste0(substr(CRASH_DATE,1,4),'-',substr(CRASH_DATE,6,7))) %>%
                  select(COLLISION_ID, CRASH_DATE, PERSON_AGE, PERSON_SEX, MONTH, TIMESTEP, yr_mo) %>% 
                  arrange(CRASH_DATE)
                )

# min(combined_df$CRASH_DATE) #"2017-01-01"
# max(combined_df$CRASH_DATE) #"2021-11-30"
```
**Filtered un-necessary data and selected columns of interest**  
<br/> 


##### Aggregate, Plot of Volume Data Available for modeling
```{r, fig.width = 10, cache=FALSE}
monthly_agg2 <- (combined_df %>% 
                   group_by(TIMESTEP, yr_mo, MONTH) %>% 
                   summarise(n = n(), .groups = 'drop') %>% 
                   arrange(TIMESTEP, yr_mo, MONTH)
                 )
plot_ly(x=monthly_agg2$yr_mo, y=monthly_agg2$n, type='bar')
```
**Notice the huge COVID-19 Impact vehicle repair volumes that's visible in this dataset!**  
<br/>  


##### Create Coarse Aggregate for modeling with feature groups
```{r, fig.width = 10, cache=FALSE}
monthly_agg3 <- (combined_df %>% 
                   group_by(TIMESTEP, yr_mo, MONTH, PERSON_SEX, PERSON_AGE) %>% 
                   summarise(n = n(), .groups = 'drop') %>% 
                   arrange(TIMESTEP, yr_mo, MONTH, PERSON_SEX, PERSON_AGE)
)
```
<br/>  


##### Create Training and Test data sets
```{r, fig.width = 10, cache=FALSE}
train_df <- monthly_agg3 %>% filter(TIMESTEP <= 47) # <= 2020-12
test_df <- monthly_agg3 %>% filter(TIMESTEP > 47) # > 2020-12
```
<br/>  


##### Train Linear Model
```{r, fig.width = 10, cache=FALSE}
lm_model <- lm(data = train_df, formula = n ~ TIMESTEP + MONTH + PERSON_AGE + PERSON_SEX ) #0.5457
summary(lm_model)
SST_lm_train <- sum((train_df$n - mean(train_df$n))^2)
SSE_lm_train <- sum((train_df$n - predict(lm_model, newdata = train_df))^2)
R_sq_lm_train <- 1-(SSE_lm_train/SST_lm_train)
R_sq_lm_train

plot_ly(x=train_df$yr_mo, y=(predict(lm_model, newdata = train_df)-train_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of Linear Model Residuals vs yr_mo for Training Data',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'LM Residuals', rangemode = "tozero"))
```
<br/>  


##### Linear Model Test Metrics
```{r, fig.width = 10, cache=FALSE}
SST_lm_test <- sum((test_df$n - mean(test_df$n))^2)
SSE_lm_test <- sum((test_df$n - predict(lm_model, newdata = test_df))^2)
R_sq_lm_test <- 1-(SSE_lm_test/SST_lm_test)
R_sq_lm_test

plot_ly(x=test_df$yr_mo, y=(predict(lm_model, newdata = test_df)-test_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of Linear Model Residuals vs yr_mo for Test Data',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'LM Residuals', rangemode = "tozero"))
```
<br/>  


##### Train Random Forest Model
```{r, fig.width = 10, cache=FALSE}
set.seed(12345)
rf_model <- randomForest(n ~ TIMESTEP + MONTH + PERSON_AGE + PERSON_SEX, data=train_df, importance=TRUE, ntree=110)  #
summary(rf_model)
var_importance_df <- data.frame(importance(rf_model)) %>% rename('PCT_IncMSE'='X.IncMSE') %>% arrange(desc(PCT_IncMSE))
var_importance_df
SST_rf_train <- sum((train_df$n - mean(train_df$n))^2)
SSE_rf_train <- sum((train_df$n - predict(rf_model, newdata = train_df))^2)
R_sq_rf_train <- 1-(SSE_rf_train/SST_rf_train)
R_sq_rf_train

plot_ly(x=seq(1,length(rf_model$mse),1), y=rf_model$mse, mode='lines+markers', type='scatter') %>% 
  layout(title = 'Plot of Random Forest Training Error vs Number of Trees',
         xaxis = list(title = 'Number of Trees'),
         yaxis = list(title = 'Error', rangemode = "tozero"))

plot_ly(x=train_df$yr_mo, y=(predict(rf_model, newdata = train_df)-train_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of Random Forest Model Residuals vs yr_mo',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'RF Residuals', rangemode = "tozero"))
```
<br/>  


##### Random Forest Model Test Metrics
```{r, fig.width = 10, cache=FALSE}
SST_rf_test <- sum((test_df$n - mean(test_df$n))^2)
SSE_rf_test <- sum((test_df$n - predict(rf_model, newdata = test_df))^2)
R_sq_rf_test <- 1-(SSE_rf_test/SST_rf_test)
R_sq_rf_test

plot_ly(x=test_df$yr_mo, y=(predict(rf_model, newdata = test_df)-test_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of Random Forest Model Residuals vs yr_mo for Test Data',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'RF Residuals', rangemode = "tozero"))
```
<br/> 


##### Train CART Model
```{r, fig.width = 10, cache=FALSE}
set.seed(12345)
min_leaf <- 3
min_split <- 3*min_leaf
cart_model <- rpart(n ~ TIMESTEP + MONTH + PERSON_AGE + PERSON_SEX, data=train_df, 
                    control = c(minsplit = min_split, minbucket = min_leaf, cp=0.01)) #default = 0.01
SST_CART_train <- sum((train_df$n - mean(train_df$n))^2)
SSE_CART_train <- sum((train_df$n - predict(cart_model, newdata = train_df))^2)
R_sq_cart_train <- 1-(SSE_CART_train/SST_CART_train)
R_sq_cart_train

data.frame(Variable_Importance = cart_model$variable.importance, 
           Variable_Importance_Pct_Tot = round(100*cart_model$variable.importance/sum(cart_model$variable.importance),0))

rpart.plot(cart_model)

plot_ly(x=train_df$yr_mo, y=(predict(cart_model, newdata = train_df)-train_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of CART Model Residuals vs yr_mo',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'CART Residuals', rangemode = "tozero"))
```
<br/>  


##### CART Model Test Metrics
```{r, fig.width = 10, cache=FALSE}
SST_cart_test <- sum((test_df$n - mean(test_df$n))^2)
SSE_cart_test <- sum((test_df$n - predict(cart_model, newdata = test_df))^2)
R_sq_cart_test <- 1-(SSE_cart_test/SST_cart_test)
R_sq_cart_test

plot_ly(x=test_df$yr_mo, y=(predict(cart_model, newdata = test_df)-test_df$n), type='scatter', mode='markers') %>% 
  layout(title = 'Plot of CART Model Residuals vs yr_mo for Test Data',
         xaxis = list(title = 'yr_mo'),
         yaxis = list(title = 'CART Residuals', rangemode = "tozero"))
```
<br/> 


##### Overall Model Summary Plots - Training Data
```{r, fig.width = 10}
plot_ly(x=train_df$n, y=(predict(cart_model, newdata = train_df)), type='scatter', mode='markers', name='CART', alpha=0.3) %>%
  add_trace(x=train_df$n, y=(predict(rf_model, newdata = train_df)), type='scatter', mode='markers', name='RF', alpha=0.3) %>%
  add_trace(x=train_df$n, y=(predict(lm_model, newdata = train_df)), type='scatter', mode='markers', name='LM', alpha=0.3) %>%
  add_trace(x=c(0,500), y=c(0,500), type='scatter', mode='line', name='Perfect', alpha=1) %>%
  layout(title = 'Plot of All Model Predictions vs Actual Values - Training Data',
         xaxis = list(title = 'Actual Value', range=c(0,700)),
         yaxis = list(title = 'Model Prediction', rangemode = "tozero"))
```
<br/>


##### Overall Model Summary Plots - Test Data
```{r, fig.width = 10}
plot_ly(x=test_df$n, y=(predict(cart_model, newdata = test_df)), type='scatter', mode='markers', name='CART', alpha=0.3) %>%
  add_trace(x=test_df$n, y=(predict(rf_model, newdata = test_df)), type='scatter', mode='markers', name='RF', alpha=0.3) %>%
  add_trace(x=test_df$n, y=(predict(lm_model, newdata = test_df)), type='scatter', mode='markers', name='LM', alpha=0.3) %>%
  add_trace(x=c(0,300), y=c(0,300), type='scatter', mode='line', name='Perfect', alpha=1) %>%
  layout(title = 'Plot of All Model Predictions vs Actual Values - Test Data',
         xaxis = list(title = 'Actual Value', range=c(0,325)),
         yaxis = list(title = 'Model Prediction', rangemode = "tozero"))
```
<br/>



##### Overall Model Performance Summary Table 
```{r, fig.width = 10, cache=FALSE}
kable(data.frame(Model_Type = c('Linear', 'Random Forest', 'CART'),
           R_sq_train = round(c(R_sq_lm_train, R_sq_rf_train, R_sq_cart_train), 2),
           R_sq_test = round(c(R_sq_lm_test, R_sq_rf_test, R_sq_cart_test), 2))) %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")%>%
  row_spec(2, background = c("yellow"))
```
**=> Moving forward with the Random Forest Model**  
<br/> 




##### Estimate Volumes Using Random Forest model
```{r, fig.width = 10, cache=FALSE}
#### Assume NYC Collision Shop has 1% of Total market share - Estimate their future monthly car volume
Shop_Market_Share <- 1/100

#Compute Actual from Test Data
test_agg_df <- test_df %>% group_by(MONTH) %>% summarise(Actual_NYC_Collisions=sum(n), Actual_Shop_Volume=round(Shop_Market_Share*sum(n), 0))

#Create dataset for future year (2021) ... maybe 2022 ...decide later with team (2021 allows comparison to real volume events)
temp1 <- data.frame(TIMESTEP = c(48:59))
temp2 <- data.frame(MONTH = c('01','02','03','04','05','06','07','08','09','10','11','12'))
temp3 <- data.frame(PERSON_AGE = c(15:100))
temp4 <- data.frame(PERSON_SEX = c('M','F'))
monthly_predictions_df <- temp1 %>% bind_cols(temp2) %>%  full_join(temp3, by=character()) %>% full_join(temp4, by=character())


#Create predictions
monthly_predictions_df$predictions <- predict(rf_model, newdata = monthly_predictions_df)
monthly_predictions_agg_df <- (monthly_predictions_df %>% 
                                 group_by(MONTH) %>% 
                                 summarise(Predicted_NYC_Collisions=round(sum(predictions), 0)) %>%
                                 mutate(Predicted_Shop_Volume = round(Shop_Market_Share*Predicted_NYC_Collisions, 0)) %>% 
                                 left_join(test_agg_df, by='MONTH') %>% 
                                 mutate(YEAR = 2021) %>% 
                                 select(YEAR, MONTH, Actual_NYC_Collisions, Actual_Shop_Volume, Predicted_NYC_Collisions, Predicted_Shop_Volume)
                                 )   
  
kable(monthly_predictions_agg_df) %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```
<br/> 


##### Plot of Actual_Shop_Volume and Predicted_Shop_Volume
```{r, fig.width = 10, cache=FALSE, warning=FALSE}
plot_ly(x=monthly_predictions_agg_df$MONTH, y=monthly_predictions_agg_df$Actual_Shop_Volume, 
        type='bar', name='Actual_Shop_Volume') %>% 
  add_trace(x=monthly_predictions_agg_df$MONTH, y=monthly_predictions_agg_df$Predicted_Shop_Volume, 
            type='bar', name='Predicted_Shop_Volume') %>% 
  layout(title = 'Plot of Actual_Shop_Volume and Predicted_Shop_Volume (Test Set)',
         xaxis = list(title = 'Months in 2021'),
         yaxis = list(title = 'Car volume'))
```
<br/> 
<br/> 
<br/> 
<br/> 


#### Additional Ideas of Interest
##### Proportion of crashes by month, gender
```{r, fig.width = 10, cache=FALSE, warning=FALSE}
kable(train_df %>% 
        group_by(PERSON_SEX) %>% 
        summarise(n = sum(n)) %>% 
    arrange(PERSON_SEX)) %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```
<br/> 


##### Proportion of crashes by month, age group
```{r, fig.width = 10, cache=FALSE, warning=FALSE}
kable(train_df %>% 
        mutate(age_group = 5*floor(PERSON_AGE/5)) %>%
        group_by(age_group) %>% 
        summarise(n = sum(n)) %>% 
        arrange(age_group)) %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```
<br/> 
<br/> 


**Use the data below to determine how much volume can be gained by an advertising campaign targeting a certain demographic**  
**<a href="https://blog.hubspot.com/agency/google-adwords-benchmark-data#:~:text=Cost%2DPer%2DAction-,Average%20Clickthrough%20Rate%20(CTR),0.35%25%20for%20a%20display%20ad.">Link</a>**  
**Across all industries, the average CTR for a search ad is 1.91%, and 0.35% for a display ad.**  
**=> Determine cost of a promotion and then outline the potential increase in marketshare for that demographic**  


